Allocation scores, WIS, and CRPS via decision theory

### 2023-08-25 alloscore_manuscript/methods_apg.tex
fix

Writing $s_{Q,\tau}(x,y) = s_{Q,C}(x,y)/L = (1-\tau)x + (y-x)_{+}$, we have
\begin{align}
s_{Q,\tau}(q_{F,\tau},y) + s_{Q,1-\tau}(q_{F,1-\tau},y) &= q_{F,\tau} + \tau(q_{F,1-\tau} - q_{F,\tau}) + (y - q_{F,\tau})_{+} + (y - q_{F,1-\tau})_{+} \\
&= y + \tau\left(q_{F,1-\tau} - q_{F,\tau} + \frac{1}{\tau}(q_{F,\tau} - y)_{+} + \frac{1}{\tau}(y - q_{F,1-\tau})_{+}\right) \\
&= y + \tau \mathrm{IS}_{2\tau}(F,y)
\end{align}
The scoring rule $\mathrm{IS}_{\alpha}(F,y)$ is the \emph{interval score} suggested by expression (43) in \cite{gneiting2007strictly} 
and formalized by \cite{bracher2021evaluating} as a means for defining the weighted interval score 
\begin{align}
\mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y)=\frac{1}{J+1 / 2}\left(w_0 |y-q_{F,.5}|+\sum_{j=1}^J w_j \operatorname{IS}_{\alpha_j}(F, y)\right).
\end{align}
So by taking $\mathcal{C} = \{(1 \pm \alpha_j)L/2 \mid j > 0\}\cup \{L/2\}$ and $p_\mathcal{C}$ symmetric across $L/2$, we get 
$S_{Q,\mathcal{C}, p_C}(F,y) = \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y)$ with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$.

@@

Consider again the resource procurement decision $x(c) \in \mathcal{X}_{cont}$ (or $\mathcal{X}_{disc}$ if $c = C_i, i \in 1,\ldots,m$) a public health worker must make in 
attempting to mitigate a pandemic across multiple locations $l \in 1,\ldots,N$.  But now we ask the worker to extend their decision to cover the contingency that the total amount of the resource $\sum x_l(C)$ that they specify for the realized cost $C$ turns out to not be available.  This could be due to demand-side factors such as a budget cut, or supply-side factors such as manufacturing or shipping failures. In either case, we assume that these factors are undetermined at time $t_0$ when decisions must be made and model them with a positive random variable $K$ giving the total amount of resource that will be available
at time $t_1$.  This adds, for each $k \in \mathbb{R}_{+}$, a component to our overall decision problem that asks the question: If the total amount $K$ available of the resourse is insufficient allow the procurements $x_l(C)$ you specify for the realized cost $C$, what alternate procurement levels $x_l^{k,C}$ would you choose? The action set for this new component is 
\begin{align}
\mathcal{X}_k = \{x \in \mathbb{R}_{+}^N \mid \sum_{l=1}^{N} x_l = k\}.
\end{align}
Note that because this decision is only relevant when $\sum x_l(C)>k$ we do not include actions which use less than $k$ total of the resource. (Why?)

In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
\begin{align}
s_{tQ}\left(x, y\right) = Ck + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
\end{align}
We see then that the cost $C$ will not effect the Bayes act for this component of the decision problem motivating the definition of a new single loss function (component), the \emph{allocation loss function}
\begin{align}
s_{A}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+},
\end{align}
we use for all $\mathcal{X}_k$.

>>>>

Writing $s_{Q,\tau}(x,y) = s_{Q,C}(x,y)/L = (1-\tau)x + (y-x)_{+}$, we have
\begin{align}
s_{Q,\tau}(q_{F,\tau},y) + s_{Q,1-\tau}(q_{F,1-\tau},y) &= q_{F,\tau} + \tau(q_{F,1-\tau} - q_{F,\tau}) + (y - q_{F,\tau})_{+} + (y - q_{F,1-\tau})_{+} \\
&= y + \tau\left(q_{F,1-\tau} - q_{F,\tau} + \frac{1}{\tau}(q_{F,\tau} - y)_{+} + \frac{1}{\tau}(y - q_{F,1-\tau})_{+}\right) \\
&= y + \tau \mathrm{IS}_{2\tau}(F,y)
\end{align}
The scoring rule $\mathrm{IS}_{\alpha}(F,y)$ is the \emph{interval score} suggested by expression (43) in \cite{gneiting2007strictly} 
and formalized by \cite{bracher2021evaluating} as a means for defining the weighted interval score 
\begin{align}
\mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y)=\frac{1}{J+1 / 2}\left(w_0 |y-q_{F,.5}|+\sum_{j=1}^J w_j \operatorname{IS}_{\alpha_j}(F, y)\right).
\end{align}
So by taking $\mathcal{C} = \{(1 \pm \alpha_j)L/2 \mid j > 0\}\cup \{L/2\}$ and $p_\mathcal{C}$ symmetric across $L/2$, we get 
$S_{Q,\mathcal{C}, p_C}(F,y) = \mathrm{WIS}_{\alpha_{\{0: J\}}}(F, y) + E_{p_{\mathcal{C}}}[y]$ with $w_j = p_{j}\alpha_j/2$ and $w_0 = p_{L/2}$.

@@

Consider again the resource procurement decision $x(c) \in \mathcal{X}_{cont}$ (or $\mathcal{X}_{disc}$ if $c = C_i, i \in 1,\ldots,m$) a public health worker must make in 
attempting to mitigate a pandemic across multiple locations $l \in 1,\ldots,N$.  But now we ask the worker to extend their decision to cover the contingency that there turns out to only be a total amount $k$ of the resource available for allocation, less than the amount $\sum x_l(C)$ that they would choose to distribute if balancing only the realized cost $C$ against potential unmet demand.  This could be due to demand-side factors such as a budget cut, or supply-side factors such as manufacturing or shipping failures. In either case, we assume that these factors are undetermined at time $t_0$ when decisions must be made and model them with a positive random variable $K$ giving the total amount of resource that will be available
at time $t_1$.  This adds, for each $k \in \mathbb{R}_{+}$, a component to our overall decision problem that asks the question: If the total amount $K$ available of the resource is insufficient allow the procurements $x_l(C)$ you specify for the realized cost $C$, what alternate procurement levels $x_l^{k,C}$ would you choose? The action set for this new component is 
\begin{align}
\mathcal{X}_k = \{x \in \mathbb{R}_{+}^N \mid \sum_{l=1}^{N} x_l = k\}.
\end{align}
Note that because this decision is only relevant when $\sum x_l(C)>k$ we do not include actions which use less than $k$ total of the resource. (Why?)

In these action set variables, the loss function we used for the unconstrained $N$ location decision problem with action set $\mathcal{X}_{cert}$ now has the expression
\begin{align}
s_{tQ}\left(x, y\right) = Ck + \sum_{l=1}^{N} L(y_l - x_l)_{+}.
\end{align}
We see then that the cost $C$ will not effect the Bayes act for this component of the decision problem motivating the definition of a new single loss function (component), the \emph{allocation loss function}
\begin{align}
s_{A}(x,y):= L\sum_{l=1}^{N} (y_l - x_l)_{+},
\end{align}
we use in deriving Bayes acts for all $\mathcal{X}_k$.
